---
title: "02_Construct_Variables"
author: "Hao Lyu"
date: "2022-11-10"
output: 
    word_document: default
    html_document: default 
---
<!--Global Chunk Settings-->
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = TRUE)
```

<!--Path and packages loaded in from 00_Master.R-->
```{r paths and packages, include=FALSE, echo=FALSE}

  # Clear the environment -----------------------------------------------------

    rm(list=ls())

  # Set working directory -----------------------------------------
      
      if (Sys.info()["user"] == "wb595473") {
  
      scriptFolder <- file.path("/Users/wb595473/OneDrive - WBG/Documents/emergency-response-procurement/04-Honduras/02-R") 
  
      projectFolder <- file.path("/Users/wb595473/WBG/DIME3 Files - FY23 MDTF project/Covid_datawork/03-Honduras/1_data")
  
      
      } else if (Sys.info()["user"] == "") {
      
      scriptFolder <- file.path("") 
      
      projectFolder <- file.path("")
      
      }


    scripts          <- file.path(scriptFolder,  "2-data_contruct"                                  ) # folder for the scripts
    raw_oncae        <- file.path(projectFolder, "1_Raw/Data_ONCAE/DCC/Final"                       ) # folder for raw data from the standard portal 
    raw_data         <- file.path(projectFolder, "1_Raw"                                            ) # folder for all raw datasets
    raw_oncae_interm <- file.path(projectFolder, "1_Raw/Data_ONCAE/DCC/Intermediate/1 - Panel data" )
    intermediate     <- file.path(projectFolder, "2_Intermediate"                                   ) # all datasets used for variable constructions 
      

  # Install and load packages -------------------------------------------------

    # Packages 

            packages <-  c(
              
              "dplyr"            ,
              "tidyr"            ,
              "haven"            ,
              "plyr"             ,
              "dplyr"            ,
              "ggplot2"          ,
              "here"             ,
              "lubridate"        , 
              "reshape2"         ,
              "scales"           ,
              "RColorBrewer"     ,
              "fmsb"             ,
              "janitor"          ,
              "stargazer"        ,
              "stringi"          ,
              "stringr"          ,
              "data.table"       ,
              "gdata"            ,
              "WDI"              

              )


      load_package <- function(pkg){
        
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        
        if (length(new.pkg))
      
      install.packages(new.pkg, dependencies = TRUE)
    
    sapply(pkg, require, character.only = TRUE)
    
    }

    load_package(packages)


```

<!--Load Datasets-->
```{r load datasets, include=FALSE, echo=FALSE}

    # aggregated data from 2018 to 2021
      # data_participants_final   <- readRDS(file.path(raw_oncae, "DATA_PARTICIPANTS.RDS"))

       data_items_final          <- readRDS(file.path(raw_oncae, "DATA_ITEMS.RDS"       ))
      
       data_contracts_final      <- readRDS(file.path(raw_oncae, "DATA_CONTRACTS.RDS"   ))
      
       data_tenders_final        <- readRDS(file.path(raw_oncae, "DATA_TENDERS.RDS"     ))

    # intermediate datasets from cleaning file 
      contract_participants <- as.data.frame(fread(file.path(intermediate,"Contract_Participants.csv"))) # 35930 tenders 
      
      tender_contract       <- as.data.frame(fread(file.path(intermediate,"Contract_Tender.csv"))) # 23622 tenders 
    
      tender_item           <- as.data.frame(fread(file.path(intermediate,"Tender_Item.csv"))) # 138873 tenders 
      
      supplier_item         <- as.data.frame(fread(file.path(intermediate,"Supplier_Item.csv"))) # 85791 tenders 
      
      data_parties_merged   <- as.data.frame(fread(file.path(intermediate,"Data_Parties_Merged.csv"))) #138873 tenders 

      
```


<!--Construct COVID Dummy-->
# COVID Dummy

### To construct COVID dummy, we search for COVID related key words in the item description, contract description, and tender description columns. The number of identification is below: 

```{r identify COVID dummy, echo=TRUE}

    # Count COVID related item, contracts, and tenders  
      covid_item <- data_items_final%>%
        mutate(covid = ifelse(str_detect(STR_ITEM_DESCRIPTION, "covid")|
                                     str_detect(STR_ITEM_DESCRIPTION, "COVID")|
                                     str_detect(STR_ITEM_DESCRIPTION, "Covid")|
                                     str_detect(STR_ITEM_DESCRIPTION, "coronavirus")|
                                     str_detect(STR_ITEM_DESCRIPTION, "coronavirus")|
                                     str_detect(STR_ITEM_DESCRIPTION, "sars-cov-2"),1,0))%>%
        mutate_at(c("covid"), ~replace_na(., 0))%>%
        group_by(covid)%>%
        dplyr::summarise(covid_item = n_distinct(ID_ITEM))%>%
        ungroup()
      
      covid_contracts <- data_contracts_final%>%
        mutate(covid = ifelse(str_detect(STR_CONTRACT_DESCRIPTION, "covid")|
                                         str_detect(STR_CONTRACT_DESCRIPTION, "COVID")|
                                         str_detect(STR_CONTRACT_DESCRIPTION, "Covid")|
                                         str_detect(STR_CONTRACT_DESCRIPTION, "coronavirus")|
                                         str_detect(STR_CONTRACT_DESCRIPTION, "coronavirus")|
                                         str_detect(STR_CONTRACT_DESCRIPTION, "sars-cov-2"),1,0))%>%
        mutate_at(c("covid"), ~replace_na(., 0))%>%
        group_by(covid)%>%
        dplyr::summarise(covid_contracts = n_distinct(ID_CONTRACT))%>%
        ungroup()
      
      covid_tenders <- data_tenders_final%>%
        mutate(covid = ifelse(str_detect(STR_TENDER_DESCRIPTION, "covid")|
                                       str_detect(STR_TENDER_DESCRIPTION, "COVID")|
                                       str_detect(STR_TENDER_DESCRIPTION, "Covid")|
                                       str_detect(STR_TENDER_DESCRIPTION, "coronavirus")|
                                       str_detect(STR_TENDER_DESCRIPTION, "coronavirus")|
                                       str_detect(STR_TENDER_DESCRIPTION, "sars-cov-2"),1,0))%>%
        mutate_at(c("covid"), ~replace_na(., 0))%>%
        group_by(covid)%>%
        dplyr::summarise(covid_tenders = n_distinct(ID))%>%
        ungroup()
      
      covid_counts <- covid_tenders%>%
        left_join(covid_contracts, by = "covid")%>%
        left_join(covid_item, by = "covid")
      
      covid_counts
    
```

### As 1 tender ID links to multiple contracts IDs and multiple item IDs, we use item, contracts, and tender descriptions together to flag COVID related tenders and generate a list of COVID tenders to define whether a product or contract related to COVID in all intermediate datasets. We have to use tender ID as it is the unique identifier in all datasets. That is to say if an item in a tender is COVID related, all the items in that tender will be flagged as COVID related, as well as the contract. 

### By this method, we identify 690 COVID related tenders
```{r generate COVID tender list, echo=TRUE}

      # flag all COVID related items, contracts, tenders in intermediate datasets 
      covid_tender_item <- tender_item%>%
        mutate(covid_item = ifelse(str_detect(STR_ITEM_DESCRIPTION, "covid")|
                                str_detect(STR_ITEM_DESCRIPTION, "COVID")|
                                str_detect(STR_ITEM_DESCRIPTION, "Covid")|
                                str_detect(STR_ITEM_DESCRIPTION, "coronavirus")|
                                str_detect(STR_ITEM_DESCRIPTION, "coronavirus")|
                                str_detect(STR_ITEM_DESCRIPTION, "sars-cov-2"),1,0),
               covid_contract = ifelse(str_detect(STR_TENDER_DESCRIPTION, "covid")|
                                str_detect(STR_TENDER_DESCRIPTION, "COVID")|
                                str_detect(STR_TENDER_DESCRIPTION, "Covid")|
                                str_detect(STR_TENDER_DESCRIPTION, "coronavirus")|
                                str_detect(STR_TENDER_DESCRIPTION, "coronavirus")|
                                str_detect(STR_TENDER_DESCRIPTION, "sars-cov-2"),1,0),
               covid = case_when(covid_item == 1 ~ 1,
                                 covid_contract == 1 ~ 1))%>%
        mutate_at(c("covid"), ~replace_na(., 0))%>%
        select(ID, covid)%>%
        distinct()
      
      covid_tender_contract <- tender_contract%>%
        mutate(covid_contract = ifelse(str_detect(STR_CONTRACT_DESCRIPTION, "covid")|
                                     str_detect(STR_CONTRACT_DESCRIPTION, "COVID")|
                                     str_detect(STR_CONTRACT_DESCRIPTION, "Covid")|
                                     str_detect(STR_CONTRACT_DESCRIPTION, "coronavirus")|
                                     str_detect(STR_CONTRACT_DESCRIPTION, "coronavirus")|
                                     str_detect(STR_CONTRACT_DESCRIPTION, "sars-cov-2"),1,0),
               covid_tender = ifelse(str_detect(STR_TENDER_DESCRIPTION, "covid")|
                                         str_detect(STR_TENDER_DESCRIPTION, "COVID")|
                                         str_detect(STR_TENDER_DESCRIPTION, "Covid")|
                                         str_detect(STR_TENDER_DESCRIPTION, "coronavirus")|
                                         str_detect(STR_TENDER_DESCRIPTION, "coronavirus")|
                                         str_detect(STR_TENDER_DESCRIPTION, "sars-cov-2"),1,0),
               covid = case_when(covid_tender == 1 ~ 1,
                                 covid_contract == 1 ~ 1))%>%
        mutate_at(c("covid"), ~replace_na(., 0))%>%
        select(ID, covid)%>%
        distinct()
      
      covid_tender <- covid_tender_item%>%
        left_join(covid_tender_contract,
                  by = "ID", suffix = c("_tender_item", "_tender_contract"))%>%
        mutate_at(c("covid_tender_contract"), ~replace_na(., 0))
        
      covid_tender <- covid_tender%>%
        mutate(covid = case_when(covid_tender_item == 1 ~ 1,
                                 covid_tender_contract == 1 ~ 1))%>%
        mutate_at(c("covid"), ~replace_na(., 0))
        
      table(covid_tender[c("covid_tender_item", "covid_tender_contract")])
      
      covid_tender <- covid_tender%>%select(ID, covid)

```


## Construct Sector Dummy
### construct 4 types of classification: Medical-COVID, Medical-NonCOVID, NonMedical-COVID, Non-Medical
### identified sector for 85422 tenders 
```{r define treatment group}

# Identify Comparison Group ------------------------------------
      
      # add COVID dummy to supplier_item data 
      supplier_item <- supplier_item%>%
        left_join(covid_tender, by = "ID")  
          # note: 474 participants that cannot link to any items i.e. ocds-lcuori-grxXEr-COT. NÂ°10-2019-1/3
      
      supplier_item <- supplier_item%>%
        mutate(covid_item = ifelse(str_detect(STR_ITEM_DESCRIPTION, "covid")|
                                         str_detect(STR_ITEM_DESCRIPTION, "COVID")|
                                         str_detect(STR_ITEM_DESCRIPTION, "Covid")|
                                         str_detect(STR_ITEM_DESCRIPTION, "coronavirus")|
                                         str_detect(STR_ITEM_DESCRIPTION, "coronavirus")|
                                         str_detect(STR_ITEM_DESCRIPTION, "sars-cov-2"),1,0),
               covid = case_when(covid == 1 ~ 1,
                                 covid_item == 1 ~ 1))%>%
        mutate_at(c("covid"), ~replace_na(., 0))
      
      # Construct and index to classify contracts to define comparison group.
      covid_item_unspsc <- supplier_item%>%
        filter(covid == 1)%>%
        select(ID, ID_PARTY, NAME_PARTY, ID_ITEM, ID_ITEM_UNSPSC, STR_ITEM_UNSPSC, STR_ITEM_DESCRIPTION, covid)
      
        # Question: should we use this list to construct 'medical' or just Segment 420000? Suggest using this list to flag medical products 
      
      # Construct Medical related dummy 
        # in UNSPSC code, 42000000 is the segment of Medical Equipment and Accessories and Supplies
        #                 41000000 is the segment of Laboratory and Measuring and Observing and Testing Equipment
        #                 51000000 is the segment of Drugs and Pharmaceutical Products
      
      # extract the first 2 charactors of unspsc and match with 42, 41, and 51  
      supplier_item$ID_ITEM_UNSPSC_SEG = substr(supplier_item$ID_ITEM_UNSPSC, start = 1, stop = 2)
      
      supplier_item <- supplier_item%>%
        mutate(medical = ifelse(ID_ITEM_UNSPSC_SEG == 42, 1,
                                ifelse(ID_ITEM_UNSPSC_SEG == 41, 1,
                                       ifelse(ID_ITEM_UNSPSC_SEG == 51, 1, 0))),
               sector = ifelse(covid == 1 & medical == 1, "Medical-COVID",
                              ifelse(covid != 1 &  medical == 1, "Medical-NonCOVID", 
                                     ifelse(covid == 1 & medical != 1, "NonMedical-COVID", "Non-Medical"))))
      
      supplier_item <- as.data.table(supplier_item)
      
      supplier_item_summary <- supplier_item[!is.na(sector),.N,by = "sector"]
      
      supplier_item_summary
      
      supplier_item <- as.data.frame(supplier_item)
      
      covid_sector <- supplier_item%>%
        # drop 474 observations that does not have item information 
        filter(!is.na(sector))%>%
        select(ID, covid, sector)%>%
        distinct()
    

```



<!--Variable Construction-->
# Task 1 

## New Winners 
### Definition: new winner means a firm that has won a contract for the first time in the survey period or has won a contract beyond 12 months. The start date of a contract denotes the contract signature date. 
### Reduce the contract-level data to month-year level

### Question: how to deal with contracts that have a signature date after the day I downloaded the data? i.e. 202212 

```{r construct new winners, echo=TRUE}

    # Identify new winners 
      
       # single out suppliers between 2018 and 2022   
        contract_supplier <- contract_participants%>%
          filter(CAT_PARTY_ROLE == "supplier"|
                   CAT_PARTY_ROLE == "supplier;tenderer"&
                   YEAR <= 2022)%>%
          arrange(ID_PARTY, DT_CONTRACT_SIGNED)
        
        # calculate time lag 
        contract_supplier <- as.data.table(contract_supplier)
        
        contract_supplier = contract_supplier[, lag_date:= shift(DT_CONTRACT_SIGNED, 1, type="lag"), by="ID_PARTY"]
        
        # calculate how many days between the 2 contracts a firm has won 
        contract_supplier$lag_date <- as_datetime(contract_supplier$lag_date, tz = lubridate::tz(contract_supplier$lag_date))
        
        contract_supplier$diff_days_contract <- interval(contract_supplier$lag_date, contract_supplier$DT_CONTRACT_SIGNED) %/% days(1)
        
        # calculate the number of months from the previous contract by the same firm 
        contract_supplier$diff_months_contract <- interval(contract_supplier$lag_date, contract_supplier$DT_CONTRACT_SIGNED) %/% months(1)
        
        # identity new winners by days 
        contract_supplier <- as.data.frame(contract_supplier)
        
        contract_supplier <- contract_supplier%>%
          mutate(new_winner = case_when(is.na(diff_days_contract) ~ 1 ,
                                        diff_days_contract > 365 ~ 1 ,
                                        TRUE ~ 0 ))
        # summary statistics by year 
        # new_winner_year <- contract_supplier%>%
        #   group_by(YEAR)%>%
        #   dplyr::summarise(num_new_winner = sum(new_winner))%>%
        #   ungroup()
        
        # summary statistics by month 
        contract_supplier$month <- strftime(contract_supplier$DT_CONTRACT_SIGNED, "%m")
        
        contract_supplier$year_month <- paste0(contract_supplier$YEAR, contract_supplier$month)
        
        new_winner_month <- contract_supplier%>%
          filter(year_month >= 201901)%>%
          group_by(year_month)%>%
          dplyr::summarise(num_new_winner = sum(new_winner))%>%
          ungroup()
        
        # # Visualization for new winners 
        # ggplot(new_winner_month, aes(x = year_month, y = num_new_winner, group = 1))+
        #   geom_line()+
        #   theme_classic()+
        #   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))
      
    # Add COVID dummy 
        contract_supplier <- left_join(contract_supplier, covid_tender,
                                       by = "ID")
        # note: 329 contracts that do not exist in tender data i.e. ocds-lcuori-MLQnAG-LPNÂº047-2019-HGS-2/6
        
        contract_supplier <- contract_supplier%>%
          mutate(covid_contract = ifelse(str_detect(STR_CONTRACT_DESCRIPTION, "covid")|
                                           str_detect(STR_CONTRACT_DESCRIPTION, "COVID")|
                                           str_detect(STR_CONTRACT_DESCRIPTION, "Covid")|
                                           str_detect(STR_CONTRACT_DESCRIPTION, "coronavirus")|
                                           str_detect(STR_CONTRACT_DESCRIPTION, "coronavirus")|
                                           str_detect(STR_CONTRACT_DESCRIPTION, "sars-cov-2"),1,0),
                 covid = case_when(covid == 1 ~ 1,
                                   covid_contract == 1 ~ 1))%>%
          mutate_at(c("covid"), ~replace_na(., 0))
  
        new_winner_month_group <- contract_supplier%>%
          filter(year_month >= 201901)%>%
          group_by(year_month, covid)%>%
          dplyr::summarise(num_new_winner = sum(new_winner))%>%
          ungroup()
        
        new_winner_month_group$covid <- replace(new_winner_month_group$covid, new_winner_month_group$covid == 0, "Non-COVID")

        new_winner_month_group$covid <- replace(new_winner_month_group$covid, new_winner_month_group$covid == 1, "COVID")
        
        ggplot(new_winner_month_group, aes(x = year_month, y = num_new_winner, group = covid, color = covid))+
          geom_line()+
          theme_classic()+
          theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
          labs(x = "Month", y = "Number of New Winners", title = "New Winners")
          
```


## Market Concentration 
### Var 1: at sector level, the number of winners per sector. 
```{r market concentration quantity}

# Market concentration (quantity) ---------------------------------
    
      # definition: the number of winners per sector 
      
      # Create month dummy
      supplier_item$month <- strftime(supplier_item$DT_CONTRACT_SIGNED, "%m")
      
      supplier_item$year <- strftime(supplier_item$DT_CONTRACT_SIGNED, "%Y")
      
      supplier_item$year_month <- paste0(supplier_item$year, supplier_item$month)
      
      supplier_item <- supplier_item%>%
        filter(year_month >= 201901 &
                 year_month <= 202208)
      
      # calculate the average number of winners per sector in each month 
      market_concentration <- supplier_item%>%
        filter(!is.na(sector))%>%
        group_by(year_month, sector)%>%
        dplyr::summarise(winner_number = n_distinct(ID_PARTY))%>%
        ungroup()
      
      
      ggplot(market_concentration, aes(x = year_month, y = winner_number, group = sector, color = sector))+
        geom_line()+
        theme_classic()+
        theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
        labs(x = "Month", y = "Number of Winners", title = "Market Concentration (Number of Winners per Sector)")
      

```


### Var 2: the share of winners per sector (sector = covid/non-covid/medical/non-medical)

```{r market concentration share}

      # change market concentration indicator from long to wide 
      market_concentration_share <- market_concentration%>%
        spread(sector, winner_number)
      
      market_concentration_share <- clean_names(market_concentration_share)
      
      market_concentration_share <- market_concentration_share%>%
        mutate_if(is.numeric, ~replace(., is.na(.), 0))%>%
        mutate(total         = medical_covid + medical_non_covid + non_medical +non_medical_covid,
               medical       = medical_covid + medical_non_covid,
               covid         = medical_covid + non_medical_covid,
               
               # the share of medical items in market 
               medical_share = medical/total,
               
               # the share of covid related items in market 
               covid_share   = covid/total,
               
               medical_covid_share = medical_covid/total)
        
      market_concentration_share_long <- market_concentration_share%>%
        select(year_month, medical_share:medical_covid_share)%>%
        gather(sector, share, medical_share:medical_covid_share, factor_key = TRUE)%>%
        mutate(sector = str_replace(sector, "medical_share", "Medical"))%>%
        mutate(sector = str_replace(sector, "covid_share", "COVID"))%>%
        mutate(sector = str_replace(sector, "medical_covid_share", "Medical-COVID"))%>%
        mutate(share = share*100)

      # plot 
      ggplot(market_concentration_share_long, aes(x = year_month, y = share, group = sector, color = sector))+
        geom_line()+
        theme_classic()+
        theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
        labs(x = "Month", y = "Market Share (%)", title = "Market Concentration (Share of Sectors in Market)")
      
      # maybe change this to proportion plot? 
    
```

## Product Classification 
### Whether a firm has increased products code during COVID (selling more than one product)? 
### Number of different products (UNSPC codes) supplied by the same firm over 6 months 
```{r product classification}
   
      # create half year dummy 
      supplier_item$month <- as.numeric(supplier_item$month)
      
      firm_item <- supplier_item%>%
        mutate(half_year = ifelse(month <= 06, "01", "02"))

      firm_item$year_half <- paste0(firm_item$year, firm_item$half_year)
      
      # calculate average UNSPSC code firms have supplied within 6 months 
      firm_item <- firm_item%>%
        group_by(ID_PARTY, year_half)%>%
        dplyr::summarise(num_unspsc = n_distinct(ID_ITEM_UNSPSC))%>%
        ungroup()
      
      firm_item_stats <- firm_item%>%
        group_by(year_half)%>%
        dplyr::summarise(num_unspsc_avg = mean(num_unspsc))%>%
        ungroup()
      
      # plot 
      ggplot(firm_item_stats, aes(x = year_half, y = num_unspsc_avg, group = 1))+
        geom_line()+
        theme_classic()+
        theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
        labs(x = "Year (every 6 months)", y = "Number of Different UNSPSC Commodity Code", title = "Change of the Number of Items")
      

```


### Number of firms that have increased the product code during COVID

```{r product classification}

      # flag firms that have increased the product code during COVID
      firm_item_year <- supplier_item%>%
        group_by(ID_PARTY, year)%>%
        dplyr::summarise(num_unspsc = n_distinct(ID_ITEM_UNSPSC))%>%
        ungroup()
      
      firm_uspsc_panel <- firm_item_year%>%
        spread(year, num_unspsc)%>%
        clean_names()%>%
        replace(is.na(.), 0)%>%
        mutate(product_increase = case_when(x2020 > x2019 | x2021 > x2019 ~ 1,
                                            TRUE ~ 0))%>%
        dplyr::rename("2019" = "x2019",
                      "2020" = "x2020",
                      "2021" = "x2021",
                      "2022" = "x2022")%>%
        dplyr::rename("ID_PARTY" = "id_party")%>%
        select(ID_PARTY, product_increase)
      
      firm_product <- left_join(supplier_item, firm_uspsc_panel,
                                by = "ID_PARTY")
      
      table(firm_product["product_increase"])

```

## Geographical Analysis - Under Construction
### Same location dummy: Whether the firm is from the same municipality, or from the same province or from the same state (we can look at different levels of geographical areas, depending on the country). [Are contracts during covid more likely to be awarded to firms from the same location as the buyer? Are bidders during covid less likely to be from the same location as the buyer?
```{r location, echo=FALSE, warning=FALSE}
     
    ## location dummy - no address, need to check original dataset 
      # Same location dummy: Whether the firm is from the same municipality, or 
      # from the same province or from the same state (we can look at different 
      # levels of geographical areas, depending on the country). [Are contracts
      # during covid more likely to be awarded to firms from the same location as 
      # the buyer? Are bidders during covid less likely to be from the same location as the buyer?
      
      supplier_buyer <- data_parties_merged%>%
        filter(CAT_PARTY_ROLE == "buyer"|
                 CAT_PARTY_ROLE == "supplier"|
                 CAT_PARTY_ROLE == "supplier;tenderer")%>%
        mutate(role_region_locality = paste0(ID, "___", ID_PARTY, "___", NAME_PARTY, "___", ADDRESS_REGION, "___", ADDRESS_LOCALITY),
               CAT_PARTY_ROLE = replace(CAT_PARTY_ROLE, CAT_PARTY_ROLE == "supplier;tenderer", "supplier"))%>%
        pivot_wider(names_from = CAT_PARTY_ROLE, 
                    values_from = role_region_locality)
      
      supplier_buyer <- data_parties_merged%>%
        filter(CAT_PARTY_ROLE == "buyer"|
                 CAT_PARTY_ROLE == "supplier"|
                 CAT_PARTY_ROLE == "supplier;tenderer")%>%
        mutate(CAT_PARTY_ROLE = replace(CAT_PARTY_ROLE, CAT_PARTY_ROLE == "supplier;tenderer", "supplier"))
      
      supplier_buyer_buyer    <- supplier_buyer%>%
        filter(CAT_PARTY_ROLE == "buyer")
      
      supplier_buyer_supplier <- supplier_buyer%>%
        filter(CAT_PARTY_ROLE == "supplier")
      
      supplier_buyer <- left_join(supplier_buyer_buyer, supplier_buyer_supplier, 
                                  by = c("ocid", "ID", "ID_PARTY"),
                                  suffix = c("_BUYER", "_SUPPLIER"))
      
      # load a county list 
      

```



# Task 2 
## Processing Time indicators 
```{r processing time, echo=TRUE, warning=FALSE} 

      tender_contract <- tender_contract%>%
        select(ID, 
               ID_CONTRACT, 
               DT_TENDER_PUB, 
               DT_TENDER_START, 
               DT_TENDER_END,
               DT_CONTRACT_SIGNED, 
               everything())%>%
        mutate(publication_start = DT_TENDER_PUB - DT_TENDER_START,
               submission_time = DT_TENDER_END - DT_TENDER_START,
               processing_time = DT_CONTRACT_SIGNED - DT_TENDER_START,
               decision_time   = DT_CONTRACT_SIGNED - DT_TENDER_END)
      
      tender_contract <- as.data.table(tender_contract)
      
      # check the overlap between contract signature, contract start and end date 
      check <- tender_contract[,.N, by = "publication_start"] # not consistent 
      
      check <- tender_contract[,.N, by = "submission_time"] # consistent: end > start 
      
      check <- tender_contract[,.N, by = "processing_time"] # consistent: signature > start 
      
      check <- tender_contract[,.N, by = "decision_time"] # consistent: signature > end 
      
      # Create month dummy
      tender_contract$month <- strftime(tender_contract$DT_CONTRACT_SIGNED, "%m")
      
      tender_contract$year <- strftime(tender_contract$DT_CONTRACT_SIGNED, "%Y")
      
      tender_contract$year_month <- paste0(tender_contract$year, tender_contract$month)
      
      tender_contract <- tender_contract%>%
        filter(year_month >= 201901 &
                 year_month <= 202208)
      
      # Add covid dummy 
      tender_contract <- left_join(tender_contract, covid_tender,
                                   by = "ID")
      
      tender_contract$covid <- replace(tender_contract$covid, tender_contract$covid == 0, "Non-COVID")
      
      tender_contract$covid <- replace(tender_contract$covid, tender_contract$covid == 1, "COVID")
      

```

### Submission time:       tender inquiry end - tender inquiry start date
```{r submission time, echo=TRUE, warning=FALSE}

      # submission time 
      submission_time <- tender_contract%>%
        group_by(year_month, covid)%>%
        dplyr::summarise(submission_time_avg = mean(submission_time))%>%
        ungroup()
      
      ggplot(submission_time, aes(x = year_month, y = submission_time_avg, group = covid, color = covid))+
        geom_line()+
        theme_classic()+
        theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
        labs(x = "Month", y = "Submission Time (days)", title = "Submission Time (Number of Days from Tender Inquiry Start to Tender Inquiry End)")


```

### Processing time: contract signature - tender inquiry start date
```{r processing time, echo=TRUE, warning=FALSE}

      # process time 
      processing_time <- tender_contract%>%
        group_by(year_month, covid)%>%
        dplyr::summarise(processing_time_avg = mean(processing_time))%>%
        ungroup()
      
      ggplot(processing_time, aes(x = year_month, y = processing_time_avg, group = covid, color = covid))+
        geom_line()+
        theme_classic()+
        theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
        labs(x = "Month", y = "Processing Time (days)", title = "Processing Time (Number of Days from Tender Inquiry Start to Contract Signature Date)")
      

```

### Decision time:  contract signature – tender inquiry end date 
```{r decision time, echo=TRUE, warning=FALSE}

      # decision time 
      decision_time <- tender_contract%>%
        group_by(year_month, covid)%>%
        dplyr::summarise(decision_time_avg = mean(decision_time))%>%
        ungroup()
      
      ggplot(decision_time, aes(x = year_month, y = decision_time_avg, group = covid, color = covid))+
        geom_line()+
        theme_classic()+
        theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+
        labs(x = "Month", y = "Decision Time (days)", title = "Decision Time (Number of Days from Tender Inquiry End to Contract Signature Date)")
      

```



# Variables that are not able to construct: 
### New Bidders: no complete bidder information in the dataset 
### The Probability a firm that had won a contract in the last X years based on the number of contracts won during COVID: no bidding information 
### SME: no SME dummy or firm size dummy from raw data 
### Supplier Main Activity: no firm registry data 
### Delivery Time: no contract deliver date 
