---
title: "Report Chile: data_tender work COVID-19"
author: "Ruggero Doino"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  pdf_document: 
    keep_tex: yes
    toc: yes
    number_sections: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage{rotating, graphicx}
   - \usepackage{dcolumn}
   - \usepackage{float}
   - \usepackage{enumitem}
   
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "/Users/ruggerodoino/Documents/GitHub/emergency-response-procurement/05-Chile/02-R") }) 
---

# data_tender Exploring

```{r Master, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# Run the Master script
source("/Users/ruggerodoino/Documents/GitHub/emergency-response-procurement/05-Chile/02-R/Master.R")

```
**E-procurement**

The e-procurement portal gathers information about `r format(nrow(data_tender), nsmall = 0, digits = 2, big.mark = ",")` tenders such as "Licitation Publica" and "Licitation Privada", from 2015 to 2021. The analysis will focus only on successful tenders ("Adjudicada"), and thus tenders that went "Cerrada"(`r format((nrow(data_tender %>% filter(tender_status == "Cerrada"))/nrow(data_tender))*100, nsmall = 0, digits = 2, big.mark = ",")` \%), "Desierta" (`r format((nrow(data_tender %>% filter(tender_status == "Desierta (o art. 3 รณ 9 Ley 19.886)"))/nrow(data_tender))*100, nsmall = 0, digits = 2, big.mark = ",")` \%), "Revocada" (`r format((nrow(data_tender %>% filter(tender_status == "Revocada"))/nrow(data_tender))*100, nsmall = 0, digits = 2, big.mark = ",")` \%), "Suspendida" (`r format((nrow(data_tender %>% filter(tender_status == "Suspendida"))/nrow(data_tender))*100, nsmall = 0, digits = 2, big.mark = ",")` \%) will be excluded. As such, the final sample will account for `r format(nrow(data_tender %>% filter(tender_status == "Adjudicada")), nsmall = 0, digits = 2, big.mark = ",")` (`r format((nrow(data_tender %>% filter(tender_status == "Adjudicada"))/nrow(data_tender))*100, nsmall = 0, digits = 2, big.mark = ",")` \% of the original sample). 

This dataset contains a wide set of variables summarized in the following list: 

\begin{itemize}
  \item Tender-level: description, title, date of publication/query/award/payment, type, buyer characteristics.
  \item Item-level: quantity, price, description, title, UNSPSC codes.
  \item Participant-level: bid details such as price and volumes, firms' name.
\end{itemize}

**Covid portal**

First, I download COVID-19 data_tender from [here](https://app.powerbi.com/view?r=eyJrIjoiNmU2NzBkNzUtYmM1Mi00NGVmLTljYWQtNTIxNTlhMTQ4ZjQ5IiwidCI6ImIwMGQ1ZjQ5LTk4YWMtNGJjNS1hMmM5LWNhZmRmNzEyMTZmMCIsImMiOjR9). The repository contains two files: "OC_COVID19.csv" (tenders related to COVID-19 emergency) and "OCItem_COVID19.csv" (list of items related to COVID-19 emergency). From now on, I will call these datasets tender-level and item-level emergency datasets, while the ordinary full sample of tenders prepared by Leandro will be called chile_compra datasets. This dataset gathers information about `r format(nrow(tender_covid), nsmall = 0, digits = 2, big.mark = ",")` tenders with the following set of variables:

\begin{itemize}
  \item Tender-level: description, title, no date of publication/query/award/payment, type, buyer characteristics.
  \item Item-level: quantity, price, only date of delivery, UNSPSC code, description. 
  \item Participant-level: no bid information.
\end{itemize}

**Merging two datasets** 

As I merge the chile_compra datasets with the tender-level emergency data_tender, I find that only `r format((nrow(tender_covid %>% filter(tender_covid_bin == "Matched"))/nrow(tender_covid))*100, nsmall = 0, digits = 2)` \% of covid emergency tenders can be identified in the chile_compra datasets. However, discrapencies could be explained by several reasons: 

\begin{itemize}
  \item Emergency portal covers also tenders from 2022, while our sample ranges from 2015 to 2021. Indeed, `r format(nrow(tender_covid %>% filter(year_2022 == 1)), nsmall = 0, digits = 2)` (`r format((nrow(tender_covid %>% filter(year_2022 == 1))/nrow(tender_covid))*100, nsmall = 2, digits = 2)` \%) are from 2022. 
  \item Emergency portal gathers information not only about "Licitation Privada" and "Licitation Publica", which are the only type of tenders present in the ordinary e-procurement portal. Indeed, the emergency portal includes `r format(nrow(tender_covid %>% filter(ClaseDeCompra == "LICITACION PUBLICA" | ClaseDeCompra == "LICITACION PRIVADA")), nsmall = 1, digits = 2)` (`r format((nrow(tender_covid %>% filter(ClaseDeCompra == "LICITACION PUBLICA" | ClaseDeCompra == "LICITACION PRIVADA"))/nrow(tender_covid))*100, nsmall = 1, digits = 2)` \%).
  \item Since we focus only on successful tenders from the ordinary e-procurement portal, `r format(nrow(tender_covid %>% filter(ClaseDeCompra != "Enviada al Proveedor" | ClaseDeCompra != "Aceptada")), nsmall = 1, digits = 2)` (`r format((nrow(tender_covid %>% filter(ClaseDeCompra == "LICITACION PUBLICA" | ClaseDeCompra == "LICITACION PRIVADA"))/nrow(tender_covid))*100, nsmall = 0, digits = 2)` \%) unsuccessful emergency tenders will not be necessary. The table below better explores this point. 
  \item `r format(nrow(tender_covid %>% filter(is.na(ID))), nsmall = 1, digits = 2)` (`r format((nrow(tender_covid %>% filter(is.na(ID)))/nrow(tender_covid))*100, nsmall = 1, digits = 2)` \%) emergency tenders do not show any tender ID. 
\end{itemize}

```{r freq table for tag, echo = FALSE, results='asis', warning = FALSE, header=FALSE}

    # Create two way freq tab to see how the unmatched tenders are distributed
    tab_1 <- tender_covid %>%                     # set the dataset to work on
      tabyl(ClaseDeCompra, tender_covid_bin)  %>%  # select the variables of interest and compute a two-way frequency table
      adorn_totals( where = c("row", "col")) %>%  # compute totals by row and cols
      adorn_percentages("all")               %>%  # add percentages to the table
      adorn_pct_formatting(digits = 0)       %>%  # format with 0 digits
      adorn_ns( position = "front" )  

names(tab_1)[1] <-"Tender Type"

tab_1 %>%
  kable(format = "latex",
        booktabs = T, 
        longtable = T, 
        caption = "Frequency table: Tender type and matching with covid-19", 
        digits = 2, format.args = list(big.mark = ",")) %>%
  kable_styling(latex_options = c("striped", "HOLD_position"))

```

The emergency portal can only partially help us to identify which tenders are covid-related since the merging rate with the main sample is only `r format((nrow(tender_covid %>% filter(tender_covid_bin == "Matched"))/nrow(tender_covid))*100, nsmall = 0, digits = 2)` \%. Therefore, I will also define tenders based on the description of the tender.

*Definition of covid-related tender* Covid-related tender is defined  as any tender that meet one of the two following conditions:
\begin{itemize}
  \item The tender has been registered in the covid emergency e-procurement portal. However, we need to keep in mind that `r format(((nrow(tender_covid %>% filter(!grepl("COVID|covid|Covid|CORONAVIRUS|covid-19", get("Descripcion OC")) & !grepl("COVID|covid|Covid|CORONAVIRUS|covid-19", get("Nombre OC")))))/(nrow(tender_covid)))*100, nsmall = 0, digits = 2)` \% of tenders do not have a string such as "COVID-19", "COVID" and/or "CORONAVIRUS" in its description.  
  \item The tender has in its description a string such as "COVID-19", "COVID" and/or "CORONAVIRUS".  
\end{itemize}

# Task 1 
 
\begin{itemize}
  \item Product classification: We define three different categories: 
  \begin{enumerate}
    \item medical products, covid related (UNSPSC code starts with 42 + the item is part of the covid emergency data\_tender).
    \item medical products, non-covid related (UNSPSC code starts with 42 + the item is not part of the covid emergency data\_tender)
    \item non medical products (UNSPSC code does not start with 42)
  \end{enumerate}
\end{itemize}


# Task 2

# Task 3
