---
title: "Report Chile: data_tender work COVID-19"
author: "Ruggero Doino"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  pdf_document: 
    keep_tex: yes
    toc: yes
    number_sections: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage{rotating, graphicx}
   - \usepackage{dcolumn}
   - \usepackage{float}
   - \usepackage{enumitem}
   
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "/Users/ruggerodoino/Documents/GitHub/emergency-response-procurement/05-Chile/02-R") }) 
---

# Explore Data

```{r Master, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# Run the Master script
source("/Users/ruggerodoino/Documents/GitHub/emergency-response-procurement/05-Chile/02-R/Master.R")

```

**E-procurement**

The e-procurement portal gathers information about `r n_tenders` tenders such as "Licitation Publica" and "Licitation Privada", from 2015 to 2021. The analysis will focus only on successful tenders ("Adjudicada"), and thus tenders that went "Cerrada" (`r p_tenders_cerrada` \%), "Desierta" (`r p_tenders_desierta` \%), "Revocada" (`r p_tenders_revocada` \%), "Suspendida" (`r p_tenders_suspendida` \%) will be excluded. As such, the final sample will account for `r p_tenders_adjudicada` (`r p_tenders_subsample` \% of the original sample). 

This dataset contains a wide set of variables summarized in the following list: 

\begin{itemize}
  \item Tender-level: description, title, date of publication/query/award/payment, type, buyer characteristics.
  \item Item-level: quantity, price, description, title, UNSPSC codes.
  \item Participant-level: bid details such as price and volumes, firms' name.
\end{itemize}

**Covid portal**

First, I download COVID-19 data_tender from [here](https://app.powerbi.com/view?r=eyJrIjoiNmU2NzBkNzUtYmM1Mi00NGVmLTljYWQtNTIxNTlhMTQ4ZjQ5IiwidCI6ImIwMGQ1ZjQ5LTk4YWMtNGJjNS1hMmM5LWNhZmRmNzEyMTZmMCIsImMiOjR9). The repository contains two files: "OC_COVID19.csv" (tenders related to COVID-19 emergency) and "OCItem_COVID19.csv" (list of items related to COVID-19 emergency). From now on, I will call these datasets tender-level and item-level emergency datasets, while the ordinary full sample of tenders prepared by Leandro will be called chile_compra datasets. This dataset gathers information about `r n_tender_covid` tenders with the following set of variables:

\begin{itemize}
  \item Tender-level: description, title, no date of publication/query/award/payment, type, buyer characteristics.
  \item Item-level: quantity, price, only date of delivery, UNSPSC code, description. 
  \item Participant-level: no bid information.
\end{itemize}

**Merging two datasets** 

As I merge the chile_compra datasets with the tender-level emergency data_tender, I find that only `r p_covid_match` \% of covid emergency tenders can be identified in the chile_compra datasets. However, discrapencies could be explained by several reasons: 

\begin{itemize}
  \item Emergency portal covers also tenders from 2022, while our sample ranges from 2015 to 2021. Indeed, `r n_tenders_2022` (`r p_tenders_2022` \%) are from 2022. 
  \item Emergency portal gathers information not only about "Licitation Privada" and "Licitation Publica", which are the only type of tenders present in the ordinary e-procurement portal. Indeed, the emergency portal includes `r p_tenders_pub_priv` (`r p_tenders_pub_priv` \%).
  \item Since we focus only on successful tenders from the ordinary e-procurement portal, `r n_tenders_success` (`r p_tenders_success` \%) unsuccessful emergency tenders will not be necessary. The table below better explores this point. 
  \item `r n_tenders_without_id` (`r p_tenders_without_id` \%) emergency tenders do not show any tender ID. 
\end{itemize}

```{r freq table for tag, echo = FALSE, results='asis', warning = FALSE, header=FALSE}

tab_1 %>%
  kable(format = "latex",
        booktabs = T, 
        longtable = T, 
        caption = "Frequency table: Tender type and matching with covid-19", 
        digits = 2, format.args = list(big.mark = ",")) %>%
  kable_styling(latex_options = c("striped", "HOLD_position"))

```

The emergency portal can only partially help us to identify which tenders are covid-related since the merging rate with the main sample is only `r p_covid_match` \%. Therefore, I will also define tenders based on the description of the tender.

# Data Cleaning

## Offer Value ($AMT\_VALUE\_AWARDED$)

\enumerate{begin}
  \item I replace `r n_negative_amount_awarded` values as missing: they are negative numbers.
  \item `r n_case_check_1` observations present the following irregularity: quantity awarded is equal to total price offered (very weird) plus the value awarded is the product of these two terms (even more weird). After checking a sample of 10 cases, I realize thatthe value awarded is equal to the price offered. I replace them.
```{r check-url-1, echo = FALSE, results='asis', warning = FALSE, header=FALSE}

check_1

```
  \item `r n_case_check_2` observation present the following irregularity: all offer values are equal to 1. I check 10 random ones, and they seem to be just mistypes. I replace them as missing. 
```{r check-url-2, echo = FALSE, results='asis', warning = FALSE, header=FALSE}

check_2

```
\enumerate{end}

# Task 1 

**Define variable: medical equipment** We can define items as medical equipment following the UNSPC code: item codes that start with "42" are considered as such. The following table shows the distribution of "Medical Equipment" among "Tender Emergency".

```{r freq table for med, echo = FALSE, results='asis', warning = FALSE, header=FALSE}

tab_2 %>%
  kable(format = "latex",
        booktabs = T, 
        longtable = T, 
        caption = "Frequency table: medical equipment among emergency tenders", 
        digits = 2, format.args = list(big.mark = ",")) %>%
  kable_styling(latex_options = c("striped", "HOLD_position"))

```

**Define variable: tender-covid** We first need to analyse the composition of emergency tenders and decide a rule to establish which share of total value can be considered 

# Task 2

# Task 3
