---
title: "Covid_Waves"
author: "Hao Lyu"
date: '2022-09-07'
output:
  word_document: default
  html_document: 
    code_folding: "hide"
    chunk_output_type: inline
---
<!--Global Chunk Settings-->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

<!--path and packages loaded in from Master.R-->
```{r paths and packages, include=FALSE}

# Cleaning the environment 
  rm(list=ls()) 


# Setting paths
  if (Sys.info()["user"] == "wb595473"){ 
    
    projectFolder     <- "/Users/wb595473/WBG/DIME3 Files - FY23 MDTF project/Covid_datawork/01-global_health_data/1_data"
    scriptFolder      <- "/Users/wb595473/OneDrive - WBG/Documents/emergency-response-procurement/02-global_health_data"

    } else if (Sys.info()["user"] == "") {
      
      projectFolder    <- "" 
      scriptFolder     <- ""
      
      }   # Maria - please enter your path here


    scripts         <- file.path(scriptFolder,  "02-R/" )    
    raw             <- file.path(projectFolder, "1_Raw//")                                             
    intermediate    <- file.path(projectFolder, "2_Intermediate//")
    cleaned         <- file.path(projectFolder, "3_Cleaned//")
    output          <- file.path(projectFolder, "4_Output//")


# Loading packages
  packages <- c(
    
    "here"                 ,
    "readxl"               ,
    "haven"                ,
    "tidyverse"            ,
    "dplyr"                ,
    "tidyr"                , 
    "data.table"           ,
    "tidyfast"             ,
    "stringi"              ,
    "pacman"               ,
    "janitor"              ,
    "devtools"             ,
    "lubridate"            ,
    "stringr"              ,
    "stargazer"            ,
    "ggplot2"              ,
    "purrr"                ,
    "kableExtra"           , 
    "parallel"             
   
     )


  load_package <- function(pkg){
    
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    
    if (length(new.pkg))
      
      install.packages(new.pkg, dependencies = TRUE)
    
    sapply(pkg, require, character.only = TRUE)
    
    }

```


<!-- Load and Construct Data-->
```{r load data, include = FALSE}

# Import Data
  covid <- read.csv(paste0(raw, "owid-covid-data.csv"), stringsAsFactors = FALSE)

# Calculate Week sum 
  
    # generate a year_week variable 
    
    covid <- as.data.frame(covid)
    
    covid <- covid %>%
      mutate(date = as.Date(date, format = "%Y-%m-%d")) %>%
      mutate(yearweek = as.integer(strftime(date, format = "%Y%V")))%>%
      mutate(yearweek2 = ifelse(test = day(date) < 7 & as.integer(substr(yearweek, 5, 6)) > 51,
                                yes =  yearweek - 100,
                                no = yearweek))%>%
      select(-yearweek)%>%
      rename(yearweek = yearweek2)
      

     # calculate weekly sum of number of death, number of cases
     
     covid$location <- as.character(covid$location)
   
     death_week <- covid%>%
       group_by(location, yearweek)%>%
       summarise(death_week = sum(new_deaths))%>%
       mutate(week = 1:n())%>%
       ungroup()
    
     cases_week <- covid%>%
       group_by(location, yearweek)%>%
       summarise(cases_week = sum(new_cases))%>%
       mutate(week = 1:n())%>%
       ungroup()
     


```

<!--Draw Plots-->
```{r plot for the number of death, fig.keep='last', include=FALSE}
  
# Plot for the number of death 
     
      death_week_brazil <- filter(death_week, location == "Brazil")

      brazil_death <- 
        ggplot(death_week_brazil, aes(x = week, y = death_week))+
        geom_line(size = 1)+
        labs(x = "Week", y = "Number of New Death")+
        theme_classic()

     death_week_honduras <- filter(death_week, location == "Honduras")
     
     honduras_death <- 
       ggplot(death_week_honduras, aes(x = week, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()
     
     death_week_chile <- filter(death_week, location == "Chile")
     
     chile_death <- 
       ggplot(death_week_chile, aes(x = week, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()
     
     death_week_romania <- filter(death_week, location == "Romania")
     
     romania_death <- 
       ggplot(death_week_romania, aes(x = week, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()

     death_week <- ggarrange(brazil_death, romania_death, chile_death, honduras_death, 
                             labels = c("Brazil","Romania","Chile", "Honduras"),
                             font.label = list(size = 10, face = "plain", color ="black"),
                             ncol = 2, nrow = 2,
                             hjust = -0.5,
                             vjust = 0.8,
                             common.legend = TRUE)
     
     annotate_figure(death_week, top = text_grob("The Number of New Deaths Per Week (2020-02-24 to 2022-08-31)", 
                                           color = "Black", face = "plain", size = 12))
     

```     

### According to the number of death per week, we propose the definition of 'covid waves' as below:
*	Brazil:
    +   1st: 202023   (Jun) – 202032 (Aug) (exceeded 7,000 deaths per week) 
    +   2nd: 202109 (Mar)   – 202126 (Jun) (exceeded 10,000 deaths per week)
    +   3rd: 202205  (Feb)  –  202207 (Feb) (exceeded 5,000 deaths per week)
    +   4th: 202225  (Jun)  –  202230 (Jul) (exceeded 1,500 deaths per week)
*	Romania:
    +   1st: 202046  (Nov)  -  202051 (Dec) (exceeded 1,000 deaths per week)
    +   2nd: 202114 (Apr)   -  202123 (Jun) (exceeded 1,000 deaths per week)
    +   3rd: 202139 (Sept)  -  202147 (Nov) (exceeded 1,000 deaths per week)
*	Chile:
    +   1st: 202024 (Jun)   -  202026 (Jun) (exceeded 1,000 deaths per week)
    +   2nd: 202104 (Jan)   -  202130 (Jul) (exceeded 500 deaths per week)
    +   3rd: 202206 (Feb)   - 202211 (Mar) (exceeded 500 deaths per week)
*	Honduras:
    +   1st: 202026 (Jun)   -  202040 (Sept) (exceeded 100 deaths per week)
    +   2nd: 202104 (Jan)   -  202141 (Oct) (exceeded 100 deaths per week)
    +   3rd: 202207 (Feb)   -  202209 (Mar) (exceeded 50 deaths per week)



```{r plot for the number of new cases, fig.keep='last', include=FALSE}
# Plot the number of cases    

     cases_week$location <- as.factor(cases_week$location)
    
     cases_week_brazil <- filter(cases_week, location == "Brazil")
     
     brazil_cases <- 
       ggplot(cases_week_brazil, aes(x = week, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week_romania <- filter(cases_week, location == "Romania")
     
     romania_cases <- 
       ggplot(cases_week_romania, aes(x = week, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week_chile <- filter(cases_week, location == "Chile")
     
     chile_cases <- 
       ggplot(cases_week_chile, aes(x = week, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week_honduras <- filter(cases_week, location == "Honduras")
     
     honduras_cases <- 
       ggplot(cases_week_honduras, aes(x = week, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week <- ggarrange(brazil_cases, romania_cases, chile_cases, honduras_cases, 
                             labels = c("Brazil","Romania","Chile", "Honduras"),
                             font.label = list(size = 10, face = "plain", color ="black"),
                             ncol = 2, nrow = 2,
                             hjust = -0.5,
                             vjust = 0.8,
                             common.legend = TRUE)
     
     annotate_figure(cases_week, top = text_grob("The Number of New Cases Per Week (2020-02-24 to 2022-08-31)", 
                                                 color = "Black", face = "plain", size = 12))
     

     
```

### According to the number of new cases per week, we propose the definition of 'covid waves' as below:
*	Brazil:
    +   1st: 202026 (Jan) - 202036 (Sept) (peaked to 200,000 new cases/week)
    +   2nd: 202109 (Mar) -  202126 (Jun) (peaked to 400,000 new cases/week)
    +   3rd: 202203 (Jan) -  202205 (Feb) (peaked to 1000,000 new cases/week)
    +   4th: 202225 (Jun) -  202228 (Jul) (peaked to 400,000 new cases/week)
*	Romania:
    +   1st: 202044 (Oct) - 202051(Dec) (peaked to 30,000 new cases/week)
    +   2nd: 202110 (Mar) - 202114 (Apr) (peaked to 30,000 new cases/week)
    +   3rd: 202139 (Sept) - 202144 (Nov)(peaked to  50,000 new cases/ week)
    +   4th: 202228 (Jul) - 202232 (Aug) (peaked to 30,000 new cases/week)
*	Chile:
    +   1st: 202021(May)  - 202025 (Jun) (peaked to 30,000 new cases/week)
    +   2nd: 202109 (Mar) - 202125 (Jun) (peaked to  30,000 new cases/ week)
    +   3rd: 202204 (Jan) - 202211 (Mar) (peaked to 100,000 new cases/week)
    +   4th: 202220 (May) - 202234 (Aug) (peaked to   30,000 new cases/week)
*	Honduras:
    +   1st: 202026 (Jun) - 202040 (Oct) (peaked to 5,000 new cases per week)
    +   2nd: 202102 (Jan) - 202138 (Sept) (peaked to 5,000 new cases per week)
    +   3rd: 202203 (Jan) - 202208 (Feb) (peaked to 5,000 new cases per week)
    +   4th: 202229 (Jul) - 202233 (Aug) (peaked to 5,000 new cases per week)

    
### I would recommend using the number of new cases per week as the indicator for 'covid waves'. We can observe clearer pattern from the data. At the same time, this number won't be affected by other endogenous variables, for instance the improvement of medical treatment. The number of death won't be a good indicator as it is said when coronavirus mutated, the likelihood of causing death became lower. Also, the lower number of death in later time period might be the result of better medical treatment as healthcare providers produced more advanced medicines, lowering the likelihood of death. The improvement of medical treatment might be correlated with our variable of interest, public procurement. As government expenditure could have boosted the invention of new drugs and expanded the distribution of medical equipment. Therefore, I would propose to use the number of new cases per week as the indicator for 'covid waves'. 


