---
title: "Covid_Waves"
author: "Hao Lyu"
date: '2022-09-01'
output:
  word_document: default
  html_document: default
---
<!--Global Chunk Settings-->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

<!--path and packages loaded in from Master.R-->
```{r paths and packages, include=FALSE}

# Cleaning the environment 
  rm(list=ls()) 


# Setting paths
  if (Sys.info()["user"] == "wb595473"){ 
    
    projectFolder     <- "/Users/wb595473/WBG/DIME3 Files - FY23 MDTF project/Covid_datawork/01-global_health_data/1_data"
    scriptFolder      <- "/Users/wb595473/OneDrive - WBG/Documents/emergency-response-procurement/02-global_health_data"

    } else if (Sys.info()["user"] == "") {
      
      projectFolder    <- "" 
      scriptFolder     <- ""
      
      }   # Maria - please enter your path here


    scripts         <- file.path(scriptFolder,  "02-R/" )    
    raw             <- file.path(projectFolder, "1_Raw//")                                             
    intermediate    <- file.path(projectFolder, "2_Intermediate//")
    cleaned         <- file.path(projectFolder, "3_Cleaned//")
    output          <- file.path(projectFolder, "4_Output//")


# Loading packages
  packages <- c(
    
    "here"                 ,
    "readxl"               ,
    "haven"                ,
    "tidyverse"            ,
    "dplyr"                ,
    "tidyr"                , 
    "data.table"           ,
    "tidyfast"             ,
    "stringi"              ,
    "pacman"               ,
    "janitor"              ,
    "devtools"             ,
    "lubridate"            ,
    "stringr"              ,
    "stargazer"            ,
    "ggplot2"              ,
    "purrr"                ,
    "kableExtra"           , 
    "parallel"             
   
     )


  load_package <- function(pkg){
    
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    
    if (length(new.pkg))
      
      install.packages(new.pkg, dependencies = TRUE)
    
    sapply(pkg, require, character.only = TRUE)
    
    }

```


<!-- Load and Construct Data-->
```{r load data, include = FALSE, eval=FALSE}

# Import Data
  covid <- fread(paste0(raw, "owid-covid-data.csv"), stringsAsFactors = FALSE)

# Calculate Week sum 
  
    # generate a year_week variable 
    
    covid <- as.data.frame(covid)
    
    covid <- covid %>%
      mutate(date = as.Date(date, format = "%Y-%m-%d")) %>%
      mutate(yearweek = as.integer(strftime(date, format = "%Y%V")))%>%
      mutate(yearweek2 = ifelse(test = day(date) < 7 & as.integer(substr(yearweek, 5, 6)) > 51,
                                yes =  yearweek - 100,
                                no = yearweek))%>%
      select(-yearweek)%>%
      rename(yearweek = yearweek2)

     # calculate weekly sum of number of death, number of cases
     
     covid$location <- as.character(covid$location)
   
     death_week <- covid%>%
       group_by(location, yearweek)%>%
       summarise(death_week = sum(new_deaths))%>%
       ungroup()
    
     cases_week <- covid%>%
       group_by(location, yearweek)%>%
       summarise(cases_week = sum(new_cases))%>%
       ungroup()
     


```

<!--Draw Plots-->
```{r plot for the number of death, fig.keep='last', eval=FALSE}
  
# Plot for the number of death 
     
     death_week_brazil <- filter(death_week, location == "Brazil")
     
     brazil_death <- 
       ggplot(death_week_brazil, aes(x = yearweek, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()
     
     death_week_honduras <- filter(death_week, location == "Honduras")
     
     honduras_death <- 
       ggplot(death_week_honduras, aes(x = yearweek, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()
     
     death_week_chile <- filter(death_week, location == "Chile")
     
     chile_death <- 
       ggplot(death_week_chile, aes(x = yearweek, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()
     
     death_week_romania <- filter(death_week, location == "Romania")
     
     romania_death <- 
       ggplot(death_week_romania, aes(x = yearweek, y = death_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Death")+
       theme_classic()

     ggarrange(brazil_death, romania_death, chile_death, honduras_death, 
               labels = c("Brazil","Romania","Chile", "Honduras"),
               font.label = list(size = 11, face = "plain", color ="black"),
               ncol = 2, nrow = 2,
               hjust = -0.5,
               vjust = 0.8,
               common.legend = TRUE)

```     

### According to the number of death per week, we propose the definition of 'co-
### vid waves' as below:
* Brazil:
    + 1st wave: 202023 - 202032 (exceeded  7,000 deaths per week)
    + 2nd wave: 202109 - 202126 (exceeded 10,000 deaths per week)
    + 3rd wave: 202205 - 202207 (exceeded  5,000 deaths per week)
    + 4th wave: 202225 - 202230 (exceeded  1,500 deaths per week)
* Romania: 
    + 1st wave: 202046 - 202051 (exceeded  1,000 deaths per week)
    + 2nd wave: 202114 - 202123 (exceeded  1,000 deaths per week)
    + 3rd wave: 202139 - 202147 (exceeded  1,000 deaths per week)
* Chile: 
    + 1st wave: 202024 - 202026 (exceeded  1,000 deaths per week)
    + 2nd wave: 202104 - 202130 (exceeded    500 deaths per week)
    + 3rd wave: 202206 - 202211 (exceeded    500 deaths per week)

* Honduras:
    + 1st wave: 202026 - 202040 (exceeded 100 deaths per week)
    + 2nd wave: 202104 - 202141 (exceeded 100 deaths per week)
    + 3rd wave: 202207 - 202209 (exceeded  50 deaths per week)


```{r plot for the number of new cases, fig.keep='last', eval=FALSE}
# Plot the number of cases    


     cases_week$location <- as.factor(cases_week$location)
    
     cases_week_brazil <- filter(cases_week, location == "Brazil")
     
     brazil_cases <- 
       ggplot(cases_week_brazil, aes(x = yearweek, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week_romania <- filter(cases_week, location == "Romania")
     
     romania_cases <- 
       ggplot(cases_week_romania, aes(x = yearweek, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week_chile <- filter(cases_week, location == "Chile")
     
     chile_cases <- 
       ggplot(cases_week_chile, aes(x = yearweek, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     cases_week_honduras <- filter(cases_week, location == "Honduras")
     
     honduras_cases <- 
       ggplot(cases_week_honduras, aes(x = yearweek, y = cases_week))+
       geom_line(size = 1)+
       labs(x = "Week", y = "Number of New Cases")+
       theme_classic()
     
     ggarrange(brazil_cases, romania_cases, chile_cases, honduras_cases, 
               labels = c("Brazil","Romania","Chile", "Honduras"),
               font.label = list(size = 11, face = "plain", color ="black"),
               ncol = 2, nrow = 2,
               hjust = -0.5,
               vjust = 0.8,
               common.legend = TRUE)
     
```

### According to the number of new cases per week, we propose the definition of 
### 'covid waves' as below:
* Brazil:
    + 1st wave: 202030 - 2020 (exceeded  300,000 new cases per week)
    + 2nd wave: 202109 - 202126 (exceeded 10,000 deaths per week)
    + 3rd wave: 202205 - 202207 (exceeded  5,000 deaths per week)
    + 4th wave: 202225 - 202230 (exceeded  1,500 deaths per week)
* Romania: 
    + 1st wave: 202046 - 202051 (exceeded  1,000 deaths per week)
    + 2nd wave: 202114 - 202123 (exceeded  1,000 deaths per week)
    + 3rd wave: 202139 - 202147 (exceeded  1,000 deaths per week)
* Chile: 
    + 1st wave: 202024 - 202026 (exceeded  1,000 deaths per week)
    + 2nd wave: 202104 - 202130 (exceeded    500 deaths per week)
    + 3rd wave: 202206 - 202211 (exceeded    500 deaths per week)

* Honduras:
    + 1st wave: 202026 - 202040 (exceeded 100 deaths per week)
    + 2nd wave: 202104 - 202141 (exceeded 100 deaths per week)
    + 3rd wave: 202207 - 202209 (exceeded  50 deaths per week)
