---
title: "Data Construct"
author: "Victor Gamarra"
date: "7/27/2022"
output: pdf_document
---


```{r}

# 1) Set working directory
#if (Sys.info()['user'] == "vgamarra" ) {   
#  setwd("C:/Users/vgamarra/Documents/GitHub/emergency-response-procurement/02-R/2-data_construct")
#} else {  
# setwd("C:/Users/wb554125/Documents/GitHub/ie-dr-endline/IE_DR_endline/1. Data cleaning")
#}
#Maria enter path to data here}

```


# Setup packages

```{r setup, include=FALSE}
packages <-
	c(
		"ggplot2",
		"tidyverse",
		"data.table",
		"dplyr",
		"readxl",
		"haven",
		"here",
		"rgdal",
		"writexl",
		"lubridate",
		"skimr",
		"googleLanguageR",
		"cld2",
		"writexl",
		"zoo",
		"rvest",
		"fuzzyjoin",
		"forecast",
		"BBmisc"
		)
  

pacman::p_load(packages,
               character.only = TRUE, # So we can list all packages in a single vector
               dependencies = TRUE)   # Install packages that are required by the packages we listed

```

# Loading raw data

```{r}

#awards.final <- readRDS("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/romania-procurement-data-analytics/DataWork/Data/Final/awards-constructed.rds")
awards.final <- readRDS("C:/Users/vgamarra/Dropbox/RA-UChicago/DIME-WB/Romania-Justicia/Dashboard/Data/Final/awards-constructed.rds")

# bids.final <- readRDS("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/romania-procurement-data-analytics/DataWork/Data/Final/bids-constructed.rds")

#complaints.final <- readRDS("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/romania-procurement-data-analytics/DataWork/Data/Final/complaints-constructed.rds")
complaints.final <- readRDS("C:/Users/vgamarra/Dropbox/RA-UChicago/DIME-WB/Romania-Justicia/Dashboard/Data/Final/complaints-constructed.rds")

#initiation.final <- readRDS("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/romania-procurement-data-analytics/DataWork/Data/Final/initiation-constructed.rds")
initiation.final <- readRDS("C:/Users/vgamarra/Dropbox/RA-UChicago/DIME-WB/Romania-Justicia/Dashboard/Data/Final/initiation-constructed.rds")

```


# Variables analysis between new and old data

```{r}
# Initiation data 22 is at the initiation-lot level, we have to collapse to initiation level 
initiation.final <- initiation.final %>% 
  distinct(initiation_notice_id, .keep_all = TRUE) %>% 
  rename(value_est_ron = value_est_procedure)

```


# Product classification

## Scrapping of the TED codes, importing the CPV codes classification and creating the 3 categories of product used for the analysis: Medical covid, Non-medical covid, and non-medical.

```{r}

# Extracting the table from the TED dataset
link <- paste0("https://simap.ted.europa.eu/web/simap/covid-related-tenders/")
webpage <- read_html(link)
data <- html_nodes(webpage,"table")
covid.TED.codes<- html_table(data[[1]],header = TRUE)

# formating the name of the CVP
covid.TED.codes <- covid.TED.codes %>% 
  separate(Description,c("Country","Description"), sep = ":") %>% 
  mutate(type_TED = "Covid related") %>% 
  distinct(Description,.keep_all = TRUE) %>% 
  select(Description, type_TED)

# importing and cleaning CPV data
CPV.data <- read_excel("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets/1-raw/cpv_codes.xlsx")
CPV.data.classified <- read_excel("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets/1-raw/cpv-product classification.xlsx")

CPV.data <- CPV.data %>% 
  mutate(Description = gsub("\\.", "", Description))
CPV.data.classified <- left_join(CPV.data, CPV.data.classified, by = "cpv_code")

# merge with the covid-19 related tenders data
covid.TED.codes <- stringdist_join(covid.TED.codes, CPV.data, by = 'Description')
covid.TED.codes <- covid.TED.codes %>% 
  select(-Description.y)

CPV.data.classified <- left_join(CPV.data.classified, covid.TED.codes, by = "cpv_code")
CPV.data.classified <- CPV.data.classified %>% 
  select(-Description.x)

# merge with the initiation data
initiation.final <- left_join(initiation.final, CPV.data.classified, by = 'cpv_code', all.x=TRUE)

# Exporting the dataset to excel
CPV.data.export <- CPV.data.classified %>% 
  mutate(type_TED = replace(type_TED,is.na(type_TED),"non_medical"),
         type_TED = replace(type_TED, type_TED=="Covid related","medical_covid"),
         type_final = type_wb,
         type_final = case_when(type_TED=="medical_covid" ~ type_TED, TRUE ~ type_final),
         type_final = ifelse(type_final=="Covid related", "medical_covid",type_final)) %>% 
  select(-c(level, D_most_disaggregated_level, description))

write_xlsx(CPV.data.export,"//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets/2-clean/cpv_codes_classified.xlsx")

# Codes for future merge with the other datasets
codes.initiation <- initiation.final %>% 
  select(initiation_notice_id, type_wb, cpv_4d)

```

## Merging the datasets to have a unique set of bids, initiations, etc. 

Here I drop the observations that have NA, or bids that are outliers (i.e., value_est_ron > 1,000'000,000), and I keep only the two more common type of procedure: *Open auction* and *Simplified procedure*. The bids that are outliers refer to 35 observations, and the two types of procedures represent the 99.2\% of the total.

```{r}

# cleaning initiation dataset
initiation.final.complete <- initiation.final %>% 
  filter(cpv_code!="NA") %>% 
  mutate(type_wb = case_when(type_TED=="Covid related" ~ type_TED, TRUE ~ type_wb),
         type_wb = ifelse(type_wb=="Covid related", "medical_covid",type_wb),
         month_yr_init = format_ISO8601(initiation_notice_date, precision = "ym"),
         init_year_number=as.numeric(initiation_year))

# cleaning awards dataset
awards.final.complete <- left_join(awards.final , codes.initiation , by  = 'initiation_notice_id')
awards.final.complete <- awards.final.complete %>% 
  filter(type_wb!="NA")
awards.final.complete <- awards.final.complete %>% 
  filter(cpv_code!="NA") %>% 
  mutate(init_year_number=as.numeric(initiation_year))
awards.final.complete <- awards.final.complete %>% 
  filter(winner_sme!="NA")

# Modifying dates variables
awards.final.complete <- awards.final.complete %>% 
  mutate(month_yr_contract = format_ISO8601(contract_date, precision = "ym"),
         month_yr_init = format_ISO8601(initiation_notice_date, precision = "ym"),
         init_year_number=as.numeric(initiation_year),
         contract_year_number = case_when(
           contract_year==2018 ~ 1,
           contract_year==2019 ~ 2,
           contract_year==2020 ~ 3,
           contract_year==2021 ~ 4)) %>% 
  filter(month_yr_contract!="NA" | month_yr_init!="NA")

# Filtering mount of tenders and keeping only two types of procedure type
awards.final.complete <- awards.final.complete %>% 
  filter(procedure_type=="Licitatie deschisa" | procedure_type=="Procedura simplificata", value_contract_ron<1000000000)

initiation.final.complete <- initiation.final.complete %>% 
  filter(procedure_type=="Licitatie deschisa" | procedure_type=="Procedura simplificata", value_est_ron<1000000000)

```

## Cleaning bids dataset (Not available for 2022)

```{r eval = FALSE, echo = FALSE}

# cleaning bids dataset
bids.final.complete <- bids.final %>% 
  filter(bidder_cui!="NA")
bids.final.complete <- left_join(bids.final.complete , codes.initiation , by  = 'initiation_notice_id')
bids.final.complete <- bids.final.complete %>% 
  filter(type_wb!="NA" | cpv_code!="NA") %>% 
  mutate(init_year_number=as.numeric(initiation_year))

# Obtaining dates from other datasets
dates.contracts <- awards.final.complete %>% 
  select(initiation_notice_id, lot_number, winner_cui, award_notice_date, contract_date) %>% 
  rename(bidder_cui = winner_cui)
bids.final.complete <- left_join(bids.final.complete, dates.contracts, by = c('initiation_notice_id', 'lot_number', 'bidder_cui'))

# obtaining initiation date
dates.initiation <- initiation.final %>% 
  select(initiation_notice_id , initiation_notice_date)

bids.final.complete <- left_join(bids.final.complete , dates.initiation , by  = 'initiation_notice_id')

bids.final.complete <- bids.final.complete %>% 
  mutate(month_yr_init = format_ISO8601(initiation_notice_date, precision = "ym"))

```


# Creating indicators of Task 1

## Indicator: Number of bidders per tender

Number of bidders is defined as the number of firms who bid within an initiation and lot. Missing values at the firm identifier (**bidder_cui**) are excluded from the count. Then the number of bidders with no previous contract is also calculated.

Data then is collapsed at the monthly level to produce a graph.

```{r eval = FALSE, echo = FALSE}

# calculating number of bidders in total
bids.final.complete <- bids.final.complete %>% 
  group_by(initiation_notice_id, lot_number) %>% 
  mutate(bidders = n())

# calculating number of bidders per tender
bids.final.complete <- bids.final.complete %>% 
  group_by(initiation_notice_id) %>% 
  mutate(bidders_tender = n())

# bidder with no previous contract 
bids.final.complete <- bids.final.complete %>% 
  group_by(bidder_cui) %>% 
  arrange(bidder_cui) %>% 
  mutate(no_contract_before = if_else(is.na(contract_date), 1, 0),
         new_bidder_nocontract = if_else(winner=="TRUE" & no_contract_before == 1, 1, 0))

# Share of bidders that have not been awarded in a year
bids.final.complete <- bids.final.complete %>% 
  group_by(initiation_notice_id) %>% 
  arrange(bidder_cui) %>% 
  mutate(share_bidders_not_awarded = mean(new_bidder_nocontract))
         

## creating monthly bidders and plots
bids.monthly <- bids.final.complete %>% 
  group_by(month_yr_init) %>% 
  summarise(N_bidders = sum(bidders),
            N_bidders_tender = mean(bidders, na.rm = TRUE),
            N_bidders_nocontractbefore = sum(new_bidder_nocontract),
            share_bidders_not_awarded=mean(share_bidders_not_awarded)) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep="")))

bids.monthly2 <- bids.monthly %>% 
  filter(date > '2018-12-31')

# Number of bidders
ggplot(bids.monthly, aes(x=date, y=N_bidders)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Total number of bidders", caption = "Number of bidders is defined as the number of firms who bid within a contract and lot in a year.") + theme(plot.caption = element_text(hjust = 0))

ggplot(bids.monthly, aes(x=date, y=N_bidders_tender)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Average number of bidders per lot", caption = "Number of bidders is defined as the number of firms who bid for a lot.") + theme(plot.caption = element_text(hjust = 0))

ggplot(bids.monthly2, aes(x=date, y=N_bidders_nocontractbefore)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Total number of bidders with no  previous award", caption = "Number of bidders is defined as the number of firms who bid within a contract and lot in a year \n that did not have a previous contract.") + theme(plot.caption = element_text(hjust = 0))

ggplot(bids.monthly2, aes(x=date, y=share_bidders_not_awarded)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Share of bidders with no previous award", caption = "Number of bidders is defined as the number of firms who bid within a contract and lot in a year \n that did not have a previous contract.") + theme(plot.caption = element_text(hjust = 0))

## Number of bidders by sector 
bids.monthly <- bids.final.complete %>% 
  group_by(month_yr_init, type_wb) %>% 
  summarise(N_bidders = sum(bidders),
            N_bidders_tender = mean(bidders, na.rm = TRUE),
            N_bidders_nocontractbefore = sum(new_bidder_nocontract), 
            share_bidders_not_awarded=mean(share_bidders_not_awarded)) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep=""))) %>% 
  filter(type_wb!="NA")

bids.monthly2 <- bids.monthly %>% 
  filter(date > '2018-12-31')

ggplot(bids.monthly, aes(x=date, y=N_bidders, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Total number of bidders, by sector", caption = "Number of bidders is defined as the number of firms who bid within an award and lot in a year.") + theme(plot.caption = element_text(hjust = 0))

ggplot(bids.monthly, aes(x=date, y=N_bidders_tender, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Average number of bidders per tender, by sector", caption = "Number of bidders is defined as the number of firms who bid within an award and lot in a year.") + theme(plot.caption = element_text(hjust = 0))

ggplot(bids.monthly2, aes(x=date, y=N_bidders_nocontractbefore, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Total number of bidders with not previous award, by sector", caption = "Number of bidders is defined as the number of firms who bid within a contract and lot in a year \n that did not have a previous contract.") + theme(plot.caption = element_text(hjust = 0))

ggplot(bids.monthly2, aes(x=date, y=share_bidders_not_awarded, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Share of bidders with no previous award", caption = "Number of bidders is defined as the number of firms who bid within a contract and lot in a year \n that did not have a previous contract.") + theme(plot.caption = element_text(hjust = 0))

```

## Indicator: New winners and New Bidders

A *new winner* is defined as a winner that has been awarded a contract, and previously has not been awarded a contract in the previous 12 months. The contract signature date allows to calculate the number of months between contracts.

An alternative *new winner* definition can be added: a winner that has been awarded a contract, and previously has not been awarded a contract in the same sector (CPV group) in the previous 12 months. This allow to relax the previous definition and to allow new winners in other sectors. 

A *new bidder* is defined as a bidder that has no bids or an existing bid older than 12 months. The initiation date allows to calculate the number of days between contracts.

The data then is aggregated at the monthly level to produce graphical representations.

### New winner

```{r}

# New Winners and share of contracts with a new winner
awards.final.complete <- awards.final.complete %>% 
  group_by(winner_cui) %>% 
  arrange(winner_cui,contract_date, lot_number) %>% 
  mutate(diff_days_contract = difftime(contract_date, lag(contract_date) , units = "days"),
         new_winner = case_when(is.na(diff_days_contract) ~ 1 ,
                                diff_days_contract > 365 ~ 1 ,
                                TRUE ~ 0 )) 

# New winner if the firm wins a contract in another sector 
awards.final.complete <- awards.final.complete %>% 
  group_by(winner_cui, cpv_group) %>% 
  arrange(winner_cui,contract_date, lot_number) %>% 
  mutate(diff_days_contract_sector = difftime(contract_date, lag(contract_date) , units = "days"),
         new_winner_sector = case_when(is.na(diff_days_contract_sector) ~ 1 ,
                                diff_days_contract > 365 ~ 1 ,
                                TRUE ~ 0 ))

# New definition of winner: difference between initiation date-contract date (pendant)

# New Winners and shares of new winners
awards.monthly <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(month_yr_contract) %>% 
  summarise(new_winner_month = sum(new_winner),
            new_winner_sector_month = sum(new_winner_sector),
            total_contracts = n()) %>% 
  mutate(date = as.Date.character(paste(month_yr_contract,"-01",sep="")),
         share_new_winner = (new_winner_month)/total_contracts,
         share_new_winner_sector = (new_winner_sector_month)/total_contracts) 

# Plots

## New bidders and new winners
ggplot(awards.monthly, aes(date)) + 
  geom_line(aes(y = new_winner_month, colour = "new winners")) + 
   ylab(" ") + xlab("Year") + labs(title = "Number of new winners", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months.") + theme(plot.caption = element_text(hjust = 0))

## share of contracts won by a new winner (no contracts more than one year)
ggplot(awards.monthly, aes(date)) + 
  geom_line(aes(y = share_new_winner, colour = "Share new winners")) + 
  geom_line(aes(y = share_new_winner_sector, colour = "Share new Winners - sector")) + ylab("") + xlab("Year") + labs(title = "Average monthly share of contracts won by a new winner", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months. New winner - sector defines a winner as an awarded firm that has no existing \n contract or an existing contract older than 12 months in the same sector (CPV code)") + theme(plot.caption = element_text(hjust = 0))


## New Winners  and share of contracts with a new winner, by sectors
awards.monthly <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(month_yr_contract, type_wb) %>% 
  summarise(new_winner_month = sum(new_winner),
            new_winner_sector_month = sum(new_winner_sector),
            total_contracts = n(),
            share_new_winner = mean(new_winner) ,
            share_new_winner_sector = mean(new_winner_sector),
            date = as.Date.character(paste(month_yr_contract,"-01",sep=""))) %>% 
  filter(type_wb!="non_medical")

# Plots

## New Winners, by sectors
ggplot(awards.monthly, aes(x=date, y=new_winner_month, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Number of new winners", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract older than 12 months.") + theme(plot.caption = element_text(hjust = 0))

## Share of contracts won by a new winner (no contracts more than one year)
ggplot(awards.monthly, aes(x=date, y=share_new_winner, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Average monthly share of contracts won by a new winner, by type", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months.")

ggplot(awards.monthly, aes(x=date, y=share_new_winner_sector, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Average monthly share of contracts won by a new winner within sectors, by type", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months.")

```

### New Bidder (Still in construction)

```{r eval = FALSE, echo = FALSE}

# New Bidders
bids.final.complete <- bids.final.complete %>% 
  group_by(bidder_cui) %>%
  arrange(bidder_cui,initiation_notice_date) %>%
  mutate(diff_days_initiation = difftime(initiation_notice_date, lag(initiation_notice_date) , units = "days"),
         new_bidder = case_when(is.na(diff_days_initiation) ~ 1 ,
                                diff_days_initiation > 365 ~ 1 ,
                                TRUE ~ 0 ))

bids.monthly <- bids.final.complete %>% 
  filter(init_year_number>8) %>% 
  group_by(month_yr_init) %>% 
  summarise(new_bidder_month = sum(new_bidder)) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep=""))) 
 

# Plots

## New bidders and new winners
bids.winners.monthly <- left_join(bids.monthly, awards.monthly, by='date')
ggplot(bids.winners.monthly, aes(date)) + 
  geom_line(aes(y = new_winner_month, colour = "new winners")) + 
  geom_line(aes(y = new_bidder_month, colour = "new bidders")) + ylab("") + xlab("Year") + labs(title = "Number of new winners and new bidders", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months. A new bidder is defined as a bidder that has no bids or an existing bid older than \n 12 months.") + theme(plot.caption = element_text(hjust = 0))

## New Bidders, by sectors
bids.monthly <- bids.final.complete %>% 
  filter(init_year_number>8) %>% 
  group_by(month_yr_init, type_wb) %>% 
  summarise(N_bidders_month = sum(bidders),
            new_bidder_month = sum(new_bidder)) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep=""))) %>%
  filter(type_wb!="non_medical")

## Plots

## New Winners, by sectors
ggplot(awards.monthly, aes(x=date, y=new_winner_month, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Number of new winners", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract older than 12 months.") + theme(plot.caption = element_text(hjust = 0))

## New Bidders, by sectors
ggplot(bids.monthly, aes(x=date, y=new_bidder_month, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Number of new bidders", caption = "A new bidder is defined as a bidder if the firm has no existing bid older than 12 months.") + theme(plot.caption = element_text(hjust = 0))

## Share of contracts won by a new winner (no contracts more than one year)
ggplot(awards.monthly, aes(x=date, y=share_new_winner, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Average monthly share of contracts won by a new winner, by type", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months.")

ggplot(awards.monthly, aes(x=date, y=share_new_winner_sector, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Average monthly share of contracts won by a new winner within sectors, by type", caption = "A new winner is defined as an awarded firm that has no existing contract or an existing contract \n older than 12 months.")

```

## Indicator: Market concentration 

Two measures of market concentration: number of bidders and winners per sector, and total value of firms' contracts per sector

Number of bidders and winners per sector.

The *number of bidders* is defined as the sum of bidders (number of firms) within a sector and year. The *number of winners* is defined as the sum of winners (number of firms) within a sector and year.


Total monetary value of the bids per sector.

The *total monetary value of firms' contracts per sector* is calculated by obtaining the total monetary value from all contracts at the sector level (three digits of CPV code), and the total monetary value of the firm's contract at the sector level. These variables are calculated at the yearly level. Then, the market share of the firm is defined as the ratio between the total value of the firm's contract divided by the total value of the sector.

Then the data is aggregated at the monthly level to produce graphical representations. As some of the firms participate in more than one sector, there are extreme values. Those outliers are dropped in a second graph to allow a better visualization of the data.

```{r}

## number of contracts per sector (3 digits of CPV codes)
number.winners <- awards.final.complete %>% 
  filter(cpv_group!="NA") %>% 
  group_by(cpv_group, cpv_group_name) %>% 
  summarise(number_winners = n()) %>% 
  arrange(-number_winners)

number.winners.table <- number.winners[1:20,]

## plot:
ggplot(data=number.winners.table, aes(x= reorder(cpv_group,-number_winners),number_winners)) +
  geom_bar(stat="identity") + ylab(" ") + xlab("CPV codes") + labs(title = "Number of contracts per sector", caption = "CPV codes: 336 - Pharmaceutical products, 331 - Medical equipments, 452 - Construction work, 150 - Food and \n beverages, 158 - Miscellaneous food products, 504 - Repair of medical products.") + theme(plot.caption = element_text(hjust = 0))


# top 5 market share in a 3 digit sector-year (market concentration)
sector.total <- awards.final.complete %>% 
  filter(cpv_group!="NA") %>%
  group_by(cpv_group, cpv_group_name, contract_year) %>% 
  summarise(total_sector=sum(value_contract_ron, na.rm = TRUE),
            contracts_sector = n())

firms.total <- awards.final.complete %>% 
  filter(cpv_group!="NA") %>%
  group_by(winner_cui, cpv_group, cpv_group_name, contract_year) %>% 
  summarise(total_firm=sum(value_contract_ron, na.rm = TRUE),
            contracts_firm = n())

firm.sector.total <- left_join(firms.total, sector.total, by = c('cpv_group', 'cpv_group_name', 'contract_year'))

firm.sector.total <- firm.sector.total %>% 
  group_by(cpv_group, cpv_group_name, contract_year) %>% 
  arrange(cpv_group, cpv_group_name, contract_year, desc(total_firm)) %>% 
  mutate(share_firm = (total_firm/total_sector)*100,
         rank_share = row_number())

firm.sector.total.top5 <- firm.sector.total %>% 
  filter(rank_share<=5) %>% 
  group_by(cpv_group, cpv_group_name, contract_year) %>% 
  summarise(total_share = sum(share_firm))


# Plot
firm.sector.total.top5%>%
  ggplot(aes(x=total_share, color=contract_year, fill=contract_year)) +
  geom_density(alpha=0.3,size=1)+ 
  scale_x_log10()+
  labs(x= "Total share", title="Average top 5 firms' market share per sector",
       caption = "The total value of firms' contracts per sector is calculated by obtaining the total value of the firm's contract \n  at the sector level (3 digits of CPV code), divided by the total value from all contracts at the sector level") + theme(plot.caption = element_text(hjust = 0))

```

## Graphs for bidders

```{r eval = FALSE, echo = FALSE}

## number of bidders per sector (3 digits of CPV codes)
number.bidders <- bids.final.complete %>% 
  filter(cpv_group!="NA") %>% 
  group_by(cpv_group, cpv_group_name) %>% 
  summarise(number_bidders = n()) %>% 
  arrange(-number_bidders)

number.bidders.table <- number.bidders[1:20,]

## Plot
ggplot(data=number.bidders.table, aes(x= reorder(cpv_group,-number_bidders),number_bidders)) +
  geom_bar(stat="identity") + ylab(" ") + xlab("CPV codes") + labs(title = "Number of bidders per sector", caption = "CPV codes: 336 - Pharmaceutical products, 331 - Medical equipments, 452 - Construction work, 158 - Miscellaneous \n food products, 150 - Food and  beverages, 343 - Vehicle parts, 713 - engeneering services.") + theme(plot.caption = element_text(hjust = 0))
ggsave("C:/Users/vgamarra/Dropbox/RA-UChicago/DIME-WB/Romania/analysis/Figures/Number_bidders_perCPV.png")
write_xlsx(number.bidders.table,"C:/Users/vgamarra/Dropbox/RA-UChicago/DIME-WB/Romania/analysis/Tables/bidders_sector.xlsx")

```

## Total monetary value, average monetary value, and total number of tenders

```{r}

# filtering and plots
initiation.final.cpv.month <- initiation.final.complete %>% 
  filter(init_year_number>1) %>% 
  group_by(month_yr_init, type_wb) %>% 
  summarise(total_value = sum(value_est_ron, na.rm = TRUE),
            avg_value = mean(value_est_ron, na.rm = TRUE),
            number_tenders = n()) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep=""))) %>% 
  filter(type_wb!="NA")

initiation.final.cpv.month2 <- initiation.final.cpv.month %>% 
  filter(type_wb!="non_medical")

## Plots

# Total value of tenders
ggplot(initiation.final.cpv.month2, aes(x = date, y = total_value, col = type_wb)) +
  geom_line() + ylab("") + xlab("Year") + labs(title = "Total contract monetary value", caption = "Total amount is calculated by adding individual contract values per type of sector") + theme(plot.caption = element_text(hjust = 0))

# Average value of tenders
ggplot(initiation.final.cpv.month, aes(x = date, y = avg_value, col = type_wb)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Average monetary contract value", caption = "Average amount is calculated by averaging individual monetary contract values per type of sector.") + theme(plot.caption = element_text(hjust = 0))

# Number of tenders
ggplot(initiation.final.cpv.month2, aes(x = date, y = number_tenders, col = type_wb)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Number of contracts", caption = "Number of contracts is the sum of the number of contracts per sector.") + theme(plot.caption = element_text(hjust = 0))


# filtering and plots for type of procedure
initiation.final.cpv.month <- initiation.final.complete %>% 
  filter(init_year_number>1) %>% 
  group_by(month_yr_init, type_wb, procedure_type) %>% 
  summarise(total_value = sum(value_est_ron, na.rm = TRUE),
            avg_value = mean(value_est_ron, na.rm = TRUE),
            number_tenders = n()) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep=""))) %>% 
  filter(type_wb!="NA")

initiation.final.cpv.month2 <- initiation.final.cpv.month %>% 
  filter(type_wb!="non_medical" & procedure_type=="Procedura simplificata")

# Number of tenders for simplified procedures only
ggplot(initiation.final.cpv.month2, aes(x = date, y = number_tenders, col = type_wb)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Number of contracts, simplified procedure only", caption = "Number of contracts is the sum of the contracts per sector.") + theme(plot.caption = element_text(hjust = 0))

```

## Variables for complaint data and translation

```{r}

# Obtaining exporting data andto translate in google translate
appeals <- complaints.final %>% 
  select(appeal_object, appeal_id) %>% 
  distinct()

procedures <- complaints.final %>% 
  select(procedure_name, appeal_id) %>% 
  distinct()

write_xlsx(appeals,"//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets/appeals.xlsx")
write_xlsx(procedures,"//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets/procedures.xlsx")

# translating variables that are translated in R
complaints.final <- complaints.final %>% 
  mutate(
    contested_en = case_when(
      contested == "Documentatie" ~ "Documentation", 
      contested == "Rezultat" ~ "Result"),
    outcome_en = case_when(
      outcome == "Admite" ~ "Admited",
      outcome == "Altele" ~ "Other",
      outcome == "Respinge" ~ "Rejected"),
    outcome_method_en = case_when(
      outcome_method == "Pe exceptii" ~ "On exceptions", 
      outcome_method == "Pe fond" ~ "Based on"))

awards.final.complete <- awards.final.complete %>% 
  mutate(
    contract_type_en = case_when (
      contract_type == "Furnizare" ~ "Supply",
      contract_type == "Lucrari" ~ "Work",
      contract_type == "Servicii" ~ "Services"),
    procedure_type_en = case_when(
      procedure_type == "Licitatie deschisa" ~ "Open auction",
      procedure_type == "Procedura simplificata" ~ "Simplified procedure"))
  
# Importing translated Data
appeals.en <- read_excel("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets//2-clean/appeals_en.xlsx")
procedures.en <- read_excel("//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets//2-clean/procedures_en.xlsx")

# Joining translated data and exporting
complaints.final <- left_join(complaints.final, appeals.en , by = 'appeal_id')
complaints.final <- left_join(complaints.final, procedures.en , by = 'appeal_id')

complaints.final.export <- complaints.final %>% 
  select(procedure_name, procedure_name_en, appellant_name, contested_en, outcome_en, outcome_method_en, appeal_id, appeal_object,  appeal_object_en)
    
write_xlsx(complaints.final.export,"//gess-fs.d.ethz.ch/home$/vgamarra/Documents/GitHub/emergency-response-procurement/06-Romania/4-datasets//2-clean/complaints_translated.xlsx")

```


# Tasks 2

## Processing time indicators

Total processing time: contract signature - tender initiation
Submission time: submission deadline – tender initiation
Decision time: award date – submission deadline
Contracting time: contract signature - award date
Delivery time: delivery date – contract signature

```{r}

awards.final.complete <- awards.final.complete %>% 
  mutate(total_processing_time = difftime(contract_date,initiation_notice_date , units = "days"), contracting_time = difftime(contract_date,award_notice_date, units = "days"))


```

## Average amount of tenders and other indicators

```{r}

# Average mount of tenders
awards.monthly <- awards.final.complete %>% 
  filter(init_year_number>1) %>% 
  group_by(month_yr_init, procedure_type_en) %>% 
  summarise(total = sum(value_contract_ron),
            avg = mean(value_contract_ron)) %>% 
  mutate(date = as.Date.character(paste(month_yr_init,"-01",sep=""))) 

ggplot(awards.monthly, aes(x=date, y=avg, col=procedure_type_en)) + geom_line() + geom_hline(yintercept=900400, linetype="dashed", color = "black") + geom_vline(xintercept = as.numeric(as.Date("2020-03-16")), linetype="dashed", color = "red", size=0.7) + geom_vline(xintercept = as.numeric(as.Date("2020-04-16")), linetype="dashed", color = "red", size=0.7) + labs(title = "Average monetary value of contracts", caption = "Average amount is calculated by averaging individual tender values per type of sector.")

awards.monthly2 <- awards.monthly %>% 
  filter(procedure_type_en=="Simplified procedure")

ggplot(awards.monthly2, aes(x=date, y=avg)) + geom_line() + geom_hline(yintercept=900400, linetype="dashed", color = "black") + geom_vline(xintercept = as.numeric(as.Date("2020-03-16")), linetype="dashed", color = "red", size=0.7) + geom_vline(xintercept = as.numeric(as.Date("2020-04-16")), linetype="dashed", color = "red", size=0.7) + labs(title = "Average monetary value of contracts - simplified procedure only", caption = "Average amount is calculated by averaging individual tender values per type of sector.")

```

## Validation of changes due to the emergengy: Number of entities that purchase a given product and share of purchases by central entities relative to the total.

```{r}

## Average monthly central purchases share
awards.month1 <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(cpv_group, month_yr_contract) %>% 
  summarise(contract_value_cpv_month_tot = sum(value_contract_ron)) 

awards.month2 <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(cpv_group, month_yr_contract, ca_central) %>% 
  summarise(contract_value_cpv_month = sum(value_contract_ron)) 

other <- left_join(awards.month1, awards.month2, by = c('cpv_group','month_yr_contract'))

other_central <- other %>% 
  filter(ca_central=="Central") %>% 
  mutate(central_share = contract_value_cpv_month/contract_value_cpv_month_tot) %>% 
  group_by(month_yr_contract) %>% 
  summarise(avg_central_shr = mean(central_share)) %>% 
  mutate(date = as.Date.character(paste(month_yr_contract,"-01",sep=""))) 

ggplot(data=other_central, aes(x=date, y=avg_central_shr)) +
  geom_bar(stat="identity") + ylab("") + xlab("Year") + labs(title = "Average monthly central purchases share", caption = "The share is defined as the total contract value of tenders at the 3 digit of CPV code purchased by central entities \n divided by the total contract value of all tenders at the 3 digit of CPV code.") + theme(plot.caption = element_text(hjust = 0))

## Average monthly central purchases share, at the sector level
awards.month1 <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(cpv_group, month_yr_contract, type_wb) %>% 
  summarise(contract_value_cpv_month_tot = sum(value_contract_ron)) 

awards.month2 <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(cpv_group, month_yr_contract, ca_central, type_wb) %>% 
  summarise(contract_value_cpv_month = sum(value_contract_ron)) 

other <- left_join(awards.month1, awards.month2, by = c('cpv_group','month_yr_contract', 'type_wb'))

other_central <- other %>% 
  filter(ca_central=="Central") %>% 
  mutate(central_share = contract_value_cpv_month/contract_value_cpv_month_tot) %>% 
  group_by(month_yr_contract, type_wb) %>% 
  summarise(avg_central_shr = mean(central_share)) %>% 
  mutate(date = as.Date.character(paste(month_yr_contract,"-01",sep=""))) 

ggplot(other_central, aes(x = date, y = avg_central_shr, col = type_wb)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Average monthly central purchases share", caption = "The share is defined as the total contract value of tenders at the 3 digit of CPV code purchased by central entities \n divided by the total contract value of all tenders at the 3 digit of CPV code.") + theme(plot.caption = element_text(hjust = 0))

```

```{r}

# filtering and plots for type of procedure
awards.final.complete.month <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  mutate(type_proc_simplified = if_else(procedure_type_en=="Simplified procedure",1,0),
         winner_firm_sme = if_else(winner_sme=="SME",1,0)) %>% 
  group_by(month_yr_contract) %>% 
  summarise(total_value = sum(value_contract_ron, na.rm = TRUE),
            avg_value = mean(value_contract_ron, na.rm = TRUE),
            number_tenders = n(),
            share_simplified = mean(type_proc_simplified),
            share_firm_sme = mean(winner_firm_sme)) %>% 
  mutate(date = as.Date.character(paste(month_yr_contract,"-01",sep="")))

awards.final.complete.month2 <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  group_by(month_yr_contract, winner_sme) %>% 
  summarise(total_value = sum(value_contract_ron, na.rm = TRUE),
            avg_value = mean(value_contract_ron, na.rm = TRUE),
            number_tenders = n()) %>% 
  mutate(date = as.Date.character(paste(month_yr_contract,"-01",sep="")))

awards.final.complete.month3 <- awards.final.complete %>% 
  filter(contract_year_number>1) %>% 
  filter(procedure_type_en == "Simplified procedure") %>% 
  group_by(month_yr_contract, winner_sme) %>% 
  summarise(total_value = sum(value_contract_ron, na.rm = TRUE),
            avg_value = mean(value_contract_ron, na.rm = TRUE),
            number_tenders = n()) %>% 
  mutate(date = as.Date.character(paste(month_yr_contract,"-01",sep="")))


# Number of tenders, by size of winner
ggplot(awards.final.complete.month2, aes(x = date, y = number_tenders, col = winner_sme)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Number of contracts, by size of winner", caption = "Number of tender is the sum of the tenders per sector.")

ggplot(awards.final.complete.month3, aes(x = date, y = number_tenders, col = winner_sme)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Number of tenders for simplified procedures only, by size of winner", caption = "Number of tender is the sum of the tenders per sector.")

# share of contracts done by simplified procedure
ggplot(awards.final.complete.month, aes(x = date, y = share_simplified)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Share of contracts signed under simplified procedure", caption = "")

ggplot(awards.final.complete.month, aes(x = date, y = share_firm_sme)) +
  geom_line()+ ylab("") + xlab("Year") + labs(title = "Share of contracts won by sme firms", caption = "")


```




